\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[table]{xcolor}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{url}
\usepackage[hidelinks,bookmarks=true,bookmarksopen=true]{hyperref}
\usepackage{siunitx}
\usepackage{parskip}

\sisetup{group-separator={,}, detect-weight=true, detect-family=true}
\renewcommand{\arraystretch}{1.15}

\definecolor{pass}{RGB}{0,140,0}
\definecolor{warn}{RGB}{200,120,0}
\definecolor{fail}{RGB}{180,0,0}
\definecolor{note}{RGB}{60,60,180}
\definecolor{accent}{HTML}{01696F}
\definecolor{ink}{HTML}{28251D}
\definecolor{muted}{HTML}{7A7974}
\definecolor{rule}{HTML}{D4D1CA}

\newcommand{\PASS}[1]{\textcolor{pass}{\textbf{#1}}}
\newcommand{\WARN}[1]{\textcolor{warn}{\textbf{#1}}}
\newcommand{\FAIL}[1]{\textcolor{fail}{\textbf{#1}}}
\newcommand{\NOTE}[1]{\textcolor{note}{\textit{#1}}}

\titleformat{\section}
  {\normalfont\large\bfseries\color{accent}}{}{0pt}{}
  [\vspace{-2pt}\textcolor{rule}{\rule{\linewidth}{0.4pt}}]

\titleformat{\subsection}
  {\normalfont\normalsize\bfseries\color{ink}}{}{0pt}{}

\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{10pt}{2pt}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyfoot[L]{\footnotesize\color{muted}use4-learning-lab --- reference run 2026-06-11}
\fancyfoot[C]{\footnotesize\color{muted}\leftmark}
\fancyfoot[R]{\footnotesize\color{muted}\thepage}

\title{\textbf{Liquidity (LIQ) --- Factor Construction Audit}\\[2pt]
\large USE4-Faithful Implementation --- \texttt{liq\_build.ipynb}\\[4pt]
\normalsize USE4 Learning Lab --- Reference Audit}
\author{Audit Date: 2026-06-11}
\date{}

\begin{document}
\maketitle
\tableofcontents
\newpage

%──────────────────────────────────────────────────────────────────────────────
\section{Purpose and Scope}

Audits the Liquidity factor:
$\mathrm{LIQ} = 0.35\,\mathrm{STOM} + 0.35\,\mathrm{STOQ} + 0.30\,\mathrm{STOA}$
(USE4 Empirical Notes, Appendix~A) --- log share turnover over one month, the
average of 3 months, and the average of 12 months.

\smallskip
\textbf{Scope.} This audit covers every pipeline stage of
\texttt{liq\_build.ipynb} from the lab's reference run of 2026-06-11: all printed diagnostics, the
validation battery, and the output parquet schema. All numerical values are
verbatim from the executed notebook. The notebook generates no figures.

\NOTE{Reference numbers: every value in this document comes verbatim from the lab's reference run of 2026-06-11 (Sharadar data through 2026-06-10). Your rebuild will differ in row counts and tail statistics with data vintage; distribution shapes, stability coefficients, and check outcomes should match. Treat persistent sign or order-of-magnitude differences as bugs in your build until proven otherwise.}


%──────────────────────────────────────────────────────────────────────────────
\section{Output Schema}

The deliverable is \path{data/out/liq_use4.parquet}, written with
\texttt{compression="zstd"}, \newline\texttt{statistics=True}.

\begin{table}[h!]
\centering
\caption{Output schema}
\begin{tabular}{llp{7.4cm}}
\toprule
\textbf{Column} & \textbf{Dtype} & \textbf{Description} \\
\midrule
\texttt{permaticker} & \texttt{Int64} & Sharadar permanent ticker ID \\
\texttt{signal\_date} & \texttt{Date} & End-of-month rebalance date \\
\texttt{in\_estu} & \texttt{Bool} & ESTU membership on \texttt{signal\_date} \\
\texttt{mcap} & \texttt{Float64} & Market cap on \texttt{signal\_date} (cap-weighting) \\
\texttt{liq} & \texttt{Float64} & Raw descriptor: $0.35\,\mathrm{STOM}+0.35\,\mathrm{STOQ}+0.30\,\mathrm{STOA}$ \\
\texttt{liq\_score} & \texttt{Float64} & Final standardised exposure (the deliverable) \\
\texttt{n\_obs} & \texttt{UInt32} & 1 if STOM and the minimum month requirements were met \\
\bottomrule
\end{tabular}
\end{table}

%──────────────────────────────────────────────────────────────────────────────
\section{Shared Infrastructure}

All factor builds consume the shared estimation universe
\path{data/out/estu_monthly.parquet} (audited separately in
\texttt{estu\_audit.tex}); the time-series builds additionally consume the
shared daily-returns artifact (audited in \texttt{daily\_panel\_audit.tex}).
On this run the factor-level view of the shared inputs was:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item Signal calendar: \textbf{330 dates} (1999-01-29 $\to$ 2026-06-10)
  \item ESTU size: mean \textbf{2{,}494}, min \textbf{1{,}791}, max \textbf{3{,}045}
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Known Deviations from PDF Spec}

This section mirrors the complete NOT-IN-PDF master list from
\texttt{liq\_spec.ipynb} (\S12.) --- every judgment call the USE4 PDFs do not
specify, with the chosen solution and its revisit trigger. Rejected alternatives
are documented in the guide; the four-part validity arguments (USE4 requirement
$\to$ deviation $\to$ solution $\to$ defence) live in the pipeline-level
register, \path{docs/deviations_register.tex}.

\begin{longtable}{@{}p{0.5cm}p{3.6cm}p{6.0cm}p{3.8cm}@{}}
\caption{NOT-IN-PDF decisions (mirrors the guide master list)}\\
\toprule
\textbf{\#} & \textbf{Decision} & \textbf{Chosen solution} & \textbf{Revisit when} \\
\midrule
\endhead
1 & Zero-volume days (V\_t = 0) & Include in sum as V\_t/S\_t = 0 & Stage 4 --- zero trading is genuine illiquidity; including it is economically correct \\[3pt]
2 & Fully-suspended month (all-zero volume) & STOM\_$\tau$ = null & Stage 4 --- null propagates cleanly; arbitrary floor creates scale distortion \\[3pt]
3 & Minimum valid windows for STOQ & 2 of 3 & Stage 4 --- 2/3 allows recent IPOs to contribute a STOQ signal \\[3pt]
4 & Minimum valid windows for STOA & 2 of 12 & Stage 4 --- 2/12 is permissive; tradeoff between coverage and signal quality \\[3pt]
5 & Composite with partial descriptors & Renormalise available weights to 1.0 & Stage 4 --- renormalisation preserves coverage for newer stocks \\[3pt]
6 & Daily shares outstanding source & Sharadar SEP \texttt{sharesbas} & Stage 1 --- SEP daily is more granular; captures intra-quarter share changes \\[3pt]
7 & Units for V\_t/S\_t computation & \texttt{volume / (sharesbas * 1e6)} & Stage 1 --- \textbf{must verify empirically} before running; see Pitfall 3 \\[3pt]
8 & Trading day calendar & Dates present in Sharadar SEP & Stage 1 --- SEP-native calendar is automatically PIT-consistent and avoids extra dependency \\[3pt]
9 & Outlier treatment & 3$\sigma$ cap-weighted trim (USE4 standard) & Stage 5 --- standard USE4 default applied to all descriptors \\[3pt]
10 & Minimum obs for composite inclusion & MIN\_OBS = 1 (any valid STOM window) & Stage 4 --- single STOM allows maximum coverage; raise if signal quality concerns emerge \\[3pt]
\bottomrule
\end{longtable}

%──────────────────────────────────────────────────────────────────────────────
\section{Stage-by-Stage Audit}

%──────────────────────────────────
\subsection{Stage 1 --- SEP turnover}

\begin{description}[style=nextline]
  \item[PDF SPEC] Share turnover from daily volume and shares outstanding.
  \item[Implementation] SEP daily volume/sharesout; monthly sums; logs of 1/3/12-month windows.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item SEP rows: \textbf{36{,}343{,}678}; \textbf{16{,}624} tickers; \textbf{6{,}921} trading days
      \item Turnover fraction (2020+, positive days): median \textbf{0.00461} (target 0.001--0.010) \PASS{unit check OK}
      \item Impossible turnover ($>1$): \textbf{20{,}710} rows (0.057\%), 2{,}221 tickers
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 2 --- Load shared ESTU}

\begin{description}[style=nextline]
  \item[PDF SPEC] USE4 ESTU = MSCI USA IMI; the project proxy is the shared \path{estu_monthly.parquet} deliverable.
  \item[Implementation] Reads \path{data/out/estu_monthly.parquet}, filtered to \texttt{signal\_date} $\geq$ 1999-01-01.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Signal calendar: \textbf{330 dates} (1999-01-29 $\to$ 2026-06-10)
      \item ESTU universe rows: \textbf{1{,}773{,}820}
      \item ESTU size: mean \textbf{2{,}494}, min \textbf{1{,}791}, max \textbf{3{,}045}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 4 --- LIQ estimator}

\begin{description}[style=nextline]
  \item[PDF SPEC] Weighted composite of log turnover at 1/3/12-month horizons.
  \item[Implementation] $\mathrm{STOM}=\ln(\Sigma_{21d})$; $\mathrm{STOQ}=\ln(\text{mean}_{3m})$; $\mathrm{STOA}=\ln(\text{mean}_{12m})$; weights 0.35/0.35/0.30.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Raw panel: \textbf{1{,}746{,}700} rows, \textbf{330} dates, \textbf{16{,}574} tickers; duplicates \textbf{0}
      \item Spot 2026-05-29: \textbf{4{,}925} tickers, median liq \textbf{$-$1.633}; range $[-9.935,\,8.593]$
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 5 --- Outlier trim (3$\sigma$)}

\begin{description}[style=nextline]
  \item[PDF SPEC] Methodology \S2.2 trim.
  \item[Implementation] CW mean $\pm\,3\cdot$EW std (ESTU bounds, applied to all).
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Clipped (ESTU): \textbf{7{,}785} / 822{,}918 = \textbf{0.95\%} (target 0.5--2\%)
      \item Bounds: lo mean $-4.494$; hi mean $0.774$ (max $1.564$)
      \item Post-trim at 2026-05-29 (ESTU, $n=2{,}761$): mean $-1.476$, std 0.729, skew $-0.057$, kurt 0.608
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 6 --- Standardisation}

\begin{description}[style=nextline]
  \item[PDF SPEC] CW mean 0 (ESTU), EW std 1 (ESTU), applied to all stocks.
  \item[Implementation] Standard shared machinery.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item CW mean: \textbf{0.000000}; EW std: \textbf{1.000000}
      \item Fraction in $[\pm3]$: all \textbf{84.3\%} / ESTU \textbf{98.9\%} (923{,}782 non-ESTU illiquid names pull the all-universe tail)
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 7 --- Save}

\begin{description}[style=nextline]
  \item[PDF SPEC] Parquet deliverable with asserted schema.
  \item[Implementation] zstd; dtype + read-back assertions.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Rows \textbf{1{,}746{,}700}; dates \textbf{330}; tickers \textbf{16{,}574}; \textbf{30.2 MB}
      \item Schema check \PASS{PASSED}; read-back \PASS{PASSED}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────────────────────────────────────────────────
\section{Validation Checks}

All checks passed on the run dated 2026-06-11. Coverage (check 3) is evaluated
on completed months only --- the final in-progress signal date is excluded
(shared convention across all factor builds).

\begin{table}[h!]
\centering
\caption{Validation results --- run date 2026-06-11}
\small
\begin{tabular}{clll}
\toprule
\textbf{Check} & \textbf{Description} & \textbf{Observed} & \textbf{Target / Status} \\
\midrule
1a & Max $|\mu_{\text{CW}}(\texttt{liq\_score})|$ & $6.57\times10^{-16}$ & $<10^{-6}$ \quad \PASS{PASS} \\
1b & Mean EW std & $1.0000$ & $1.0\pm0.02$ \quad \PASS{PASS} \\
2 & liq in $[-6,2]$ for ESTU & $100.0\%$ $[-5.37,\,1.56]$ & $\geq95\%$ \quad \PASS{PASS} \\
3 & Min coverage post-2005 (completed months) & $4{,}636$ & $\geq4{,}000$ \quad \PASS{PASS} \\
4 & MoM Spearman $\rho$ mean & $0.9706$ & $>0.95$ \quad \PASS{PASS} \\
5 & Q1$\to$Q5 mean liq monotone & spread $2.3483$ & monotone \quad \PASS{PASS} \\
6 & Disk vs memory max diff & $0.00\times10^{0}$ & $<10^{-10}$ \quad \PASS{PASS} \\
\bottomrule
\end{tabular}
\end{table}

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item \textbf{Check 5 quintile means:} Q1 $=-2.9928$, Q2 $=-2.2370$, Q3 $=-1.8509$, Q4 $=-1.4397$, Q5 $=-0.6445$ (log turnover is negative by construction).
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Spot Check at 2026-05-29}

Aggregate diagnostics at the penultimate signal date:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item Coverage: \textbf{4{,}925} tickers; median liq \textbf{$-$1.633}
  \item Post-trim ESTU ($n=2{,}761$): mean \textbf{$-$1.476}, std \textbf{0.729}, skew \textbf{$-$0.057}, excess kurtosis \textbf{0.608}
\end{itemize}

\NOTE{No per-ticker factor values were printed by the build notebook
; the spot check reports aggregate statistics only.}

\end{document}
