\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[table]{xcolor}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{url}
\usepackage[hidelinks,bookmarks=true,bookmarksopen=true]{hyperref}
\usepackage{siunitx}
\usepackage{parskip}

\sisetup{group-separator={,}, detect-weight=true, detect-family=true}
\renewcommand{\arraystretch}{1.15}

\definecolor{pass}{RGB}{0,140,0}
\definecolor{warn}{RGB}{200,120,0}
\definecolor{fail}{RGB}{180,0,0}
\definecolor{note}{RGB}{60,60,180}
\definecolor{accent}{HTML}{01696F}
\definecolor{ink}{HTML}{28251D}
\definecolor{muted}{HTML}{7A7974}
\definecolor{rule}{HTML}{D4D1CA}

\newcommand{\PASS}[1]{\textcolor{pass}{\textbf{#1}}}
\newcommand{\WARN}[1]{\textcolor{warn}{\textbf{#1}}}
\newcommand{\FAIL}[1]{\textcolor{fail}{\textbf{#1}}}
\newcommand{\NOTE}[1]{\textcolor{note}{\textit{#1}}}

\titleformat{\section}
  {\normalfont\large\bfseries\color{accent}}{}{0pt}{}
  [\vspace{-2pt}\textcolor{rule}{\rule{\linewidth}{0.4pt}}]

\titleformat{\subsection}
  {\normalfont\normalsize\bfseries\color{ink}}{}{0pt}{}

\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{10pt}{2pt}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyfoot[L]{\footnotesize\color{muted}use4-learning-lab --- reference run 2026-06-11}
\fancyfoot[C]{\footnotesize\color{muted}\leftmark}
\fancyfoot[R]{\footnotesize\color{muted}\thepage}

\title{\textbf{Industry Factors (INDUSTRIES) --- Factor Construction Audit}\\[2pt]
\large USE4-Faithful Implementation --- \texttt{industries\_build.ipynb}\\[4pt]
\normalsize USE4 Learning Lab --- Reference Audit}
\author{Audit Date: 2026-06-11}
\date{}

\begin{document}
\maketitle
\tableofcontents
\newpage

%──────────────────────────────────────────────────────────────────────────────
\section{Purpose and Scope}

Audits the industry-factor exposure layer: USE4 specifies 60 GICS-based industry
factors with single 0/1 membership for most stocks (fractional for diversified
firms via business-segment data). With GICS and segment data both absent from
public Sharadar, the implementation delivers \textbf{55 factors} over a
hand-engineered scheme (\path{industry_scheme.csv}: 154 Sharadar
\texttt{industry} atoms $\to$ 55 factors $+$ 1 fallback) with single membership
throughout. Industries are dummies, not standardized descriptors --- there is no
trim, no standardisation, and the validation battery is membership-shaped.

\smallskip
\textbf{Scope.} This audit covers every pipeline stage of
\texttt{industries\_build.ipynb} as executed on 2026-06-11: all printed
diagnostics, the 8-check validation battery, and the output parquet schema.
All numerical values are verbatim from the executed notebook. The notebook
generates no figures.

\NOTE{Reference numbers: every value in this document comes verbatim from the lab's reference run of 2026-06-11 (Sharadar data through 2026-06-10). Your rebuild will differ in row counts with data vintage; the structural results (55 factors present every date, zero floor violations, spot checks) should match exactly.}


%──────────────────────────────────────────────────────────────────────────────
\section{Output Schema}

The deliverable is \path{data/out/industries_use4.parquet},
written with \texttt{compression="zstd"}, \newline\texttt{statistics=True}.

\begin{table}[h!]
\centering
\caption{Output schema}
\begin{tabular}{llp{7.4cm}}
\toprule
\textbf{Column} & \textbf{Dtype} & \textbf{Description} \\
\midrule
\texttt{permaticker} & \texttt{Int64} & Sharadar permanent ticker ID \\
\texttt{signal\_date} & \texttt{Date} & End-of-month rebalance date \\
\texttt{in\_estu} & \texttt{Bool} & ESTU membership on \texttt{signal\_date} \\
\texttt{mcap} & \texttt{Float64} & Market cap on \texttt{signal\_date} \\
\texttt{sharadar\_industry} & \texttt{String} & Raw classification atom (audit column) \\
\texttt{industry} & \texttt{String} & \textbf{The factor label} --- one of 55 scheme factors \\
\texttt{use4\_sector} & \texttt{String} & Sector grouping of the factor \\
\bottomrule
\end{tabular}
\end{table}

\smallskip
A second artifact, \path{data/out/industry_weights_use4.parquet}, carries the
per-date industry cap-weight vector --- the $w$ in USE4's identification
constraint $\sum_j w_j f_j^{\mathrm{ind}} = 0$:

\begin{table}[h!]
\centering
\caption{Weights artifact schema}
\begin{tabular}{llp{7.0cm}}
\toprule
\textbf{Column} & \textbf{Dtype} & \textbf{Description} \\
\midrule
\texttt{signal\_date} & \texttt{Date} & End-of-month rebalance date \\
\texttt{industry} & \texttt{String} & Factor label \\
\texttt{n\_members} & \texttt{UInt32} & In-ESTU members on the date \\
\texttt{cap\_weight} & \texttt{Float64} & Industry share of in-ESTU cap (sums to 1 per date) \\
\bottomrule
\end{tabular}
\end{table}

%──────────────────────────────────────────────────────────────────────────────
\section{Shared Infrastructure}

Consumes the shared estimation universe \path{data/out/estu_monthly.parquet}
(audited in \texttt{estu\_audit.tex}); independent of the daily-returns
artifact. On this run:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item Signal calendar: \textbf{330 dates} (1999-01-29 $\to$ 2026-06-10)
  \item Universe rows: \textbf{1{,}773{,}820}; ESTU size mean \textbf{2{,}494},
    min \textbf{1{,}791}, max \textbf{3{,}045}
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Known Deviations from PDF Spec}

This section mirrors the complete master list from
\texttt{industries\_spec.ipynb} (\S9). The four-part validity
arguments will join the pipeline-level register
(\path{docs/deviations_register.tex}) at its next revision.

\begin{longtable}{@{}p{0.5cm}p{3.6cm}p{6.0cm}p{3.8cm}@{}}
\caption{NOT-IN-PDF / NOT-IN-DATA decisions}\\
\toprule
\textbf{\#} & \textbf{Decision} & \textbf{Chosen solution} & \textbf{Revisit when} \\
\midrule
\endhead
1 & GICS unavailable \WARN{Material} & Sharadar \texttt{industry} atoms (154) + hand-engineered 55-factor scheme targeting the USE4 60-list & GICS or a crosswalk is licensed \\[3pt]
2 & No segment data \WARN{Material} & Single 0/1 membership for all stocks (USE4: $\sim$63\% of cap multi-exposed) & Segment-level data procured \\[3pt]
3 & Classification not PIT & Static current-snapshot labels & EDGAR 10-K SIC PIT layer if drift shown to matter \\[3pt]
4 & Floor rule & median $\geq$ 15 and min $\geq$ 8 ESTU members over the full sample & CSR explanatory-power screening replaces member counts \\[3pt]
5 & Floor exceptions (4) & Internet \& Catalog Retail (AMZN); Managed Health Care (USE4-explicit); Airlines; Industrial Conglomerates & Any falls below $\sim$5 members persistently \\[3pt]
6 & Merges vs the 60-list & Drilling$\to$Equip\&Svcs; Paper$+$Forest$\to$Construction Materials; both metals$\to$Metals \& Mining; durables/apparel/leisure and retail consolidated; telecom 2$\to$1 & CSR favors a split AND members support it \\[3pt]
7 & Splits vs the 60-list & Diversified Financials and Software each split in two (2026 tectonics: 107--151-member monsters as single factors) & Never --- deliberate modernisation \\[3pt]
8 & Fat factors kept whole & Banks, Real Estate, Biotech \& Life Sciences, Software--Application per USE4-2011 & CSR screening adjudicates \\[3pt]
9 & Judgment atoms & Solar$\to$Semi Equipment; Consumer Electronics (incl.\ AAPL)$\to$Computers \& Electronics; SPACs$\to$Consumer Finance \& DFS & Reclassification evidence \\[3pt]
10 & REITs & One Real Estate factor under Financials (USE4-2011, pre-2016 GICS promotion) & Modern-GICS baseline adopted \\
\bottomrule
\end{longtable}

%──────────────────────────────────────────────────────────────────────────────
\section{Stage-by-Stage Audit}

%──────────────────────────────────
\subsection{Stage 1 --- Scheme + classification load}

\begin{description}[style=nextline]
  \item[PDF SPEC] GICS-based classification.
  \item[Implementation] \path{industry_scheme.csv} loaded and asserted against the
    observed atom universe --- an unmapped atom fails the build (vendor taxonomy
    changes must be reviewed, never silently bucketed).
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Scheme: \textbf{154} atoms $\to$ \textbf{55} factors ($+$ fallback)
      \item Tickers: \textbf{44{,}163} permatickers (deduplicated)
      \item Atom coverage: all 154 observed atoms mapped \PASS{PASS}
    \end{itemize}
  \item[Deviation] \NOTE{The fail-loudly assert fired on first execution: three
    atoms (Banks--Global, Drug Manufacturers--Major, Silver) exist in the full
    ticker universe but never appeared among ESTU members, so the ESTU-derived
    design worksheet missed them. They were reviewed and added to the scheme
    (Banks, Pharmaceuticals, Metals \& Mining respectively) --- the guard working
    exactly as specified.}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 2 --- Load shared ESTU}

\begin{description}[style=nextline]
  \item[PDF SPEC] Exposures over the estimation universe.
  \item[Implementation] Shared \path{estu_monthly.parquet}, signal dates
    $\geq$ 1999-01-01.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item \textbf{330 dates}; \textbf{1{,}773{,}820} universe rows;
        ESTU mean \textbf{2{,}494}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stages 3--4 --- Classification and panel}

\begin{description}[style=nextline]
  \item[PDF SPEC] One primary industry exposure of 1.0 (fractional via segments
    for diversified firms).
  \item[Implementation] permaticker $\to$ atom $\to$ factor join; null atoms
    $\to$ \texttt{UNASSIGNED}; single membership throughout (segment data
    absent --- headline deviation \#2).
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Panel: \textbf{1{,}773{,}820} rows, \textbf{330} dates,
        \textbf{16{,}829} tickers; duplicates \textbf{0}
      \item Max in-ESTU \texttt{UNASSIGNED} on any date: \textbf{1}
      \item Spot 2026-05-29: \textbf{2{,}761} ESTU members across
        \textbf{55} factors; largest: Biotechnology \& Life Sciences
        \textbf{277}, Banks \textbf{225}, Real Estate \textbf{172},
        HC Equipment \& Technology \textbf{115}, Software--Application \textbf{107}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 5 --- Save}

\begin{description}[style=nextline]
  \item[Implementation] zstd parquet; 7-column dtype assertion; read-back assertion.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Rows \textbf{1{,}773{,}820}; dates \textbf{330}
        (1999-01-29 $\to$ 2026-06-10); tickers \textbf{16{,}829}; \textbf{8.6 MB}
      \item Schema check \PASS{PASSED}; read-back \PASS{PASSED}
      \item Weights artifact: \textbf{18{,}150} rows (330 dates $\times$ 55
        factors); per-date weight sums $|\Sigma w - 1|$ max
        $\mathbf{5.55\times10^{-16}}$; largest single weight ever
        \textbf{14.1\%}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────────────────────────────────────────────────
\section{Validation Checks}

All 8 checks passed on the run dated 2026-06-11. The battery is
membership-shaped (industries are dummies); the completed-months coverage
convention does not apply because the map is static.

\begin{table}[h!]
\centering
\caption{Validation results --- run date 2026-06-11}
\footnotesize
\begin{tabular}{@{}c@{\hspace{6pt}}p{4.6cm}p{4.6cm}p{3.6cm}@{}}
\toprule
\textbf{Check} & \textbf{Description} & \textbf{Observed} & \textbf{Target / Status} \\
\midrule
1 & In-ESTU \texttt{UNASSIGNED} per date & max $1$ & $\leq 2$ \quad \PASS{PASS} \\
2 & All 55 factors present every date & min $55$, max $55$ & $55$ \quad \PASS{PASS} \\
3 & Floor (median $\geq15$, min $\geq8$, ex-exceptions) & $0$ violations & $0$ \quad \PASS{PASS} \\
4 & Static map (one label per permaticker) & $0$ violators & $0$ \quad \PASS{PASS} \\
5 & Single membership (no duplicate rows) & $0$ dupes & $0$ \quad \PASS{PASS} \\
6 & Largest factor cap share & $14.1\%$ & $\leq 30\%$ \quad \PASS{PASS} \\
7 & Ticker spot checks & all 6 correct & \PASS{PASS} \\
8 & Disk vs memory & row-exact & \PASS{PASS} \\
\bottomrule
\end{tabular}
\end{table}

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item \textbf{Check 3 exceptions} (declared, kept on cap-weight / USE4-explicitness
    grounds): Airlines (med 12, min 9); Industrial Conglomerates (med 10, min 6);
    Internet \& Catalog Retail (med 6, min 3); Managed Health Care (med 14, min 9).
  \item \textbf{Check 7:} AAPL $\to$ Computers \& Electronics; JPM $\to$ Banks;
    XOM $\to$ Oil, Gas \& Consumable Fuels; AMZN $\to$ Internet \& Catalog Retail;
    NVDA $\to$ Semiconductors; PLD $\to$ Real Estate.
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Spot Check at 2026-05-29}

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item \textbf{2{,}761} ESTU members classified across \textbf{55} factors
  \item Five largest factors by membership: Biotechnology \& Life Sciences
    \textbf{277}, Banks \textbf{225}, Real Estate \textbf{172}, Health Care
    Equipment \& Technology \textbf{115}, Software--Application \textbf{107}
  \item Largest factor cap share at any date in the sample: \textbf{14.1\%}
\end{itemize}

\NOTE{Membership counts per factor are the spot statistics for a dummy layer;
no per-ticker exposures exist beyond the six named spot checks above.}

\end{document}
