\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[table]{xcolor}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{url}
\usepackage[hidelinks,bookmarks=true,bookmarksopen=true]{hyperref}
\usepackage{siunitx}
\usepackage{parskip}

\sisetup{group-separator={,}, detect-weight=true, detect-family=true}
\renewcommand{\arraystretch}{1.15}

\definecolor{pass}{RGB}{0,140,0}
\definecolor{warn}{RGB}{200,120,0}
\definecolor{fail}{RGB}{180,0,0}
\definecolor{note}{RGB}{60,60,180}
\definecolor{accent}{HTML}{01696F}
\definecolor{ink}{HTML}{28251D}
\definecolor{muted}{HTML}{7A7974}
\definecolor{rule}{HTML}{D4D1CA}

\newcommand{\PASS}[1]{\textcolor{pass}{\textbf{#1}}}
\newcommand{\WARN}[1]{\textcolor{warn}{\textbf{#1}}}
\newcommand{\FAIL}[1]{\textcolor{fail}{\textbf{#1}}}
\newcommand{\NOTE}[1]{\textcolor{note}{\textit{#1}}}

\titleformat{\section}
  {\normalfont\large\bfseries\color{accent}}{}{0pt}{}
  [\vspace{-2pt}\textcolor{rule}{\rule{\linewidth}{0.4pt}}]

\titleformat{\subsection}
  {\normalfont\normalsize\bfseries\color{ink}}{}{0pt}{}

\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{10pt}{2pt}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyfoot[L]{\footnotesize\color{muted}use4-learning-lab --- reference run 2026-06-11}
\fancyfoot[C]{\footnotesize\color{muted}\leftmark}
\fancyfoot[R]{\footnotesize\color{muted}\thepage}

\title{\textbf{Size (SIZE) --- Factor Construction Audit}\\[2pt]
\large USE4-Faithful Implementation --- \texttt{size\_build.ipynb}\\[4pt]
\normalsize USE4 Learning Lab --- Reference Audit}
\author{Audit Date: 2026-06-11}
\date{}

\begin{document}
\maketitle
\tableofcontents
\newpage

%──────────────────────────────────────────────────────────────────────────────
\section{Purpose and Scope}

Audits the Size factor: $\mathrm{SIZE} = 1.0 \cdot \mathrm{LNCAP}$,
$\texttt{lncap}_{i,t} = \ln(\text{mcap}_{i,t})$ (USE4 Empirical Notes, Appendix~A,
p.~51), standardised to CW mean 0 / EW std 1 within ESTU. The deliverable also
feeds NLS (\texttt{nls\_build.ipynb} reads \path{size_use4.parquet}).

\smallskip
\textbf{Scope.} This audit covers every pipeline stage of
\texttt{size\_build.ipynb} from the lab's reference run of 2026-06-11: all printed diagnostics, the
validation battery, and the output parquet schema. All numerical values are
verbatim from the executed notebook. The notebook generates no figures.

\NOTE{Reference numbers: every value in this document comes verbatim from the lab's reference run of 2026-06-11 (Sharadar data through 2026-06-10). Your rebuild will differ in row counts and tail statistics with data vintage; distribution shapes, stability coefficients, and check outcomes should match. Treat persistent sign or order-of-magnitude differences as bugs in your build until proven otherwise.}


%──────────────────────────────────────────────────────────────────────────────
\section{Output Schema}

The deliverable is \path{data/out/size_use4.parquet}, written with
\texttt{compression="zstd"}, \newline\texttt{statistics=True}.

\begin{table}[h!]
\centering
\caption{Output schema}
\begin{tabular}{llp{7.4cm}}
\toprule
\textbf{Column} & \textbf{Dtype} & \textbf{Description} \\
\midrule
\texttt{permaticker} & \texttt{Int64} & Sharadar permanent ticker ID \\
\texttt{signal\_date} & \texttt{Date} & End-of-month rebalance date \\
\texttt{in\_estu} & \texttt{Bool} & ESTU membership on \texttt{signal\_date} \\
\texttt{mcap} & \texttt{Float64} & Market cap on \texttt{signal\_date} (cap-weighting) \\
\texttt{lncap} & \texttt{Float64} & Raw descriptor: $\ln(\text{mcap})$, post 4$\sigma$ trim \\
\texttt{size\_score} & \texttt{Float64} & Final standardised exposure (the deliverable) \\
\texttt{n\_obs} & \texttt{UInt32} & 1 if a valid mcap snapshot existed \\
\bottomrule
\end{tabular}
\end{table}

%──────────────────────────────────────────────────────────────────────────────
\section{Shared Infrastructure}

All factor builds consume the shared estimation universe
\path{data/out/estu_monthly.parquet} (audited separately in
\texttt{estu\_audit.tex}); the time-series builds additionally consume the
shared daily-returns artifact (audited in \texttt{daily\_panel\_audit.tex}).
On this run the factor-level view of the shared inputs was:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item Signal calendar: \textbf{330 dates} (1999-01-29 $\to$ 2026-06-10)
  \item ESTU size: mean \textbf{2{,}494}, min \textbf{1{,}791}, max \textbf{3{,}045}
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Known Deviations from PDF Spec}

This section mirrors the complete NOT-IN-PDF master list from
\texttt{size\_spec.ipynb} (\S12.) --- every judgment call the USE4 PDFs do not
specify, with the chosen solution and its revisit trigger. Rejected alternatives
are documented in the guide; the four-part validity arguments (USE4 requirement
$\to$ deviation $\to$ solution $\to$ defence) live in the pipeline-level
register, \path{docs/deviations_register.tex}.

\begin{longtable}{@{}p{0.5cm}p{3.6cm}p{6.0cm}p{3.8cm}@{}}
\caption{NOT-IN-PDF decisions (mirrors the guide master list)}\\
\toprule
\textbf{\#} & \textbf{Decision} & \textbf{Chosen solution} & \textbf{Revisit when} \\
\midrule
\endhead
1 & Market cap source & \texttt{marketcap $\times$ 1e6} from SEP prices parquet & Stage 1 --- data load \\[3pt]
2 & Price column (if computing directly) & \texttt{closeunadj} (unadjusted, matches how Sharadar computes \texttt{marketcap}) & Stage 1 --- derivation \\[3pt]
3 & Non-positive market cap & Mark \texttt{lncap} null; exclude from all computation & Stage 1 filter \\[3pt]
4 & Calendar start & 1999-01-01 & Stage 0 parameter \\[3pt]
5 & MIN\_OBS & 1 (any valid spot obs suffices) & Stage 0 parameter \\[3pt]
6 & n\_obs semantics & Always 1 for valid rows; kept for schema consistency & Stage 3 output schema \\[3pt]
7 & Trim center (CW vs EW mean) & \textbf{Equal-weighted mean} --- CW mean ($\approx$ 25.5) is dominated by large-caps, placing the 3$\sigma$ lower bound inside the small-cap ESTU body and clipping $\sim$20\% of rows & Stage 5 \\[3pt]
8 & Trim $\sigma$ multiplier & \textbf{4$\sigma$} --- lncap is approximately Gaussian (excess kurtosis $\approx$ --0.1); a 3$\sigma$ EW upper bound clips the large-cap Q5 tail and compresses the quintile spread below the 4.0 log-unit validation target & Stage 5 \\[3pt]
\bottomrule
\end{longtable}

%──────────────────────────────────────────────────────────────────────────────
\section{Stage-by-Stage Audit}

%──────────────────────────────────
\subsection{Stage 1 --- SEP marketcap snapshots}

\begin{description}[style=nextline]
  \item[PDF SPEC] ``Given by the logarithm of the total market capitalization of the firm.''
  \item[Implementation] Loads SEP marketcap; \texttt{lncap} $= \ln(\text{mcap in dollars})$.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item SEP rows: \textbf{36{,}195{,}114}; \textbf{16{,}551} tickers (1999-01-04 $\to$ 2026-06-10)
      \item marketcap (\$M): min 0.100, median 422.9, max 10{,}121{,}939
      \item lncap range $[11.51,\,29.95]$; median \textbf{19.86}
      \item Unit spot check: $\ln(1000 \times 10^{6}) = 20.7233$ \PASS{exact}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 2 --- Load shared ESTU}

\begin{description}[style=nextline]
  \item[PDF SPEC] USE4 ESTU = MSCI USA IMI; the project proxy is the shared \path{estu_monthly.parquet} deliverable.
  \item[Implementation] Reads \path{data/out/estu_monthly.parquet}, filtered to \texttt{signal\_date} $\geq$ 1999-01-01.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Signal calendar: \textbf{330 dates} (1999-01-29 $\to$ 2026-06-10)
      \item ESTU universe rows: \textbf{1{,}773{,}820}
      \item ESTU size: mean \textbf{2{,}494}, min \textbf{1{,}791}, max \textbf{3{,}045}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 4 --- SIZE estimator (PIT join)}

\begin{description}[style=nextline]
  \item[PDF SPEC] Last observed market cap as of the signal date.
  \item[Implementation] \texttt{join\_asof} (backward) of SEP snapshots onto \texttt{estu\_monthly}.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Raw panel: \textbf{1{,}773{,}820} rows, \textbf{330} dates, \textbf{16{,}829} tickers; duplicates \textbf{0}
      \item Spot 2026-05-29: 4{,}819 tickers, \textbf{4{,}800} with lncap; ESTU median lncap \textbf{21.874}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 5 --- Outlier trim (4$\sigma$, EW mean)}

\begin{description}[style=nextline]
  \item[PDF SPEC] Methodology Notes \S2.2 trims outliers; width/centre for lncap is a documented deviation.
  \item[Implementation] Per-date EW mean $\pm\,4\sigma$ bounds from ESTU, clip applied to all stocks.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Clipped (ESTU): \textbf{404} / 822{,}918 = \textbf{0.05\%} (target $<$ 0.3\% for 4$\sigma$)
      \item Bounds: lo mean 15.886 (min 15.249); hi mean 27.162 (max 28.851)
      \item Post-trim at 2026-05-29 (ESTU, $n=2{,}761$): mean 22.063, std 1.695, skew 0.708, kurt 0.246
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 6 --- Standardisation}

\begin{description}[style=nextline]
  \item[PDF SPEC] CW mean 0 (ESTU), EW std 1 (ESTU), applied to all stocks.
  \item[Implementation] Standard shared machinery; sigma guard $>10^{-8}$.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item CW mean: \textbf{$-$0.000000}; EW std: \textbf{1.000000}
      \item Fraction in $[\pm3]$: all \textbf{44.4\%} / ESTU \textbf{80.4\%} (structural: 911{,}683 non-ESTU micro-caps sit far below the CW mean; median ESTU score $-$2.250)
    \end{itemize}
  \item[Deviation] \NOTE{The low in-band fractions are intrinsic to SIZE: the CW mean is set by mega-caps, so the bulk of names standardise deeply negative. This is the documented cross-sectional signature of LNCAP, not a defect.}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 7 --- Save}

\begin{description}[style=nextline]
  \item[PDF SPEC] Parquet deliverable with asserted schema.
  \item[Implementation] zstd; 7-column dtype assertion; read-back shape/order assertion.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Rows \textbf{1{,}773{,}820}; dates \textbf{330}; tickers \textbf{16{,}829}; \textbf{25.8 MB}
      \item Schema check \PASS{PASSED}; read-back \PASS{PASSED}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────────────────────────────────────────────────
\section{Validation Checks}

All checks passed on the run dated 2026-06-11. Coverage (check 3) is evaluated
on completed months only --- the final in-progress signal date is excluded
(shared convention across all factor builds).

\begin{table}[h!]
\centering
\caption{Validation results --- run date 2026-06-11}
\small
\begin{tabular}{clll}
\toprule
\textbf{Check} & \textbf{Description} & \textbf{Observed} & \textbf{Target / Status} \\
\midrule
1a & Max $|\mu_{\text{CW}}(\texttt{size\_score})|$ & $3.37\times10^{-15}$ & $<10^{-6}$ \quad \PASS{PASS} \\
1b & Mean EW std & $1.0000$ & $1.0\pm0.02$ \quad \PASS{PASS} \\
2a & ESTU lncap range & $[19.49,\,28.85]$ & $[16.0,\,30.0]$ \quad \PASS{PASS} \\
2b & ESTU median lncap & $21.291$ & $[21.0,\,25.0]$ \quad \PASS{PASS} \\
3 & Min coverage post-2005 (completed months) & $4{,}578$ & $\geq4{,}000$ \quad \PASS{PASS} \\
4 & MoM Spearman $\rho$ mean & $0.9967$ & $>0.99$ \quad \PASS{PASS} \\
5 & Q1$\to$Q5 lncap monotone, spread $\geq3.75$ & spread $3.8866$ & monotone \quad \PASS{PASS} \\
6 & Disk vs memory max diff & $0.00\times10^{0}$ & $<10^{-10}$ \quad \PASS{PASS} \\
\bottomrule
\end{tabular}
\end{table}

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item \textbf{Check 5 quintile means:} Q1 $=19.8945$, Q2 $=20.6055$, Q3 $=21.3155$, Q4 $=22.1411$, Q5 $=23.7810$.
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Spot Check at 2026-05-29}

Aggregate diagnostics at the penultimate signal date:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item \textbf{4{,}800} tickers with valid lncap; \textbf{2{,}761} in ESTU
  \item ESTU median lncap \textbf{21.874}; range $[19.52,\,29.26]$
  \item Post-trim ESTU: mean \textbf{22.063}, std \textbf{1.695}, skew \textbf{0.708}, excess kurtosis \textbf{0.246}
\end{itemize}

\NOTE{No per-ticker factor values were printed by the build notebook
; the spot check reports aggregate statistics only.}

\end{document}
