\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[table]{xcolor}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{url}
\usepackage[hidelinks,bookmarks=true,bookmarksopen=true]{hyperref}
\usepackage{siunitx}
\usepackage{parskip}

\sisetup{group-separator={,}, detect-weight=true, detect-family=true}
\renewcommand{\arraystretch}{1.15}

\definecolor{pass}{RGB}{0,140,0}
\definecolor{warn}{RGB}{200,120,0}
\definecolor{fail}{RGB}{180,0,0}
\definecolor{note}{RGB}{60,60,180}
\definecolor{accent}{HTML}{01696F}
\definecolor{ink}{HTML}{28251D}
\definecolor{muted}{HTML}{7A7974}
\definecolor{rule}{HTML}{D4D1CA}

\newcommand{\PASS}[1]{\textcolor{pass}{\textbf{#1}}}
\newcommand{\WARN}[1]{\textcolor{warn}{\textbf{#1}}}
\newcommand{\FAIL}[1]{\textcolor{fail}{\textbf{#1}}}
\newcommand{\NOTE}[1]{\textcolor{note}{\textit{#1}}}

\titleformat{\section}
  {\normalfont\large\bfseries\color{accent}}{}{0pt}{}
  [\vspace{-2pt}\textcolor{rule}{\rule{\linewidth}{0.4pt}}]

\titleformat{\subsection}
  {\normalfont\normalsize\bfseries\color{ink}}{}{0pt}{}

\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{10pt}{2pt}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyfoot[L]{\footnotesize\color{muted}use4-learning-lab --- reference run 2026-06-11}
\fancyfoot[C]{\footnotesize\color{muted}\leftmark}
\fancyfoot[R]{\footnotesize\color{muted}\thepage}

\title{\textbf{Beta (BETA) --- Factor Construction Audit}\\[2pt]
\large USE4-Faithful Implementation --- \texttt{beta\_build.ipynb}\\[4pt]
\normalsize USE4 Learning Lab --- Reference Audit}
\author{Audit Date: 2026-06-11}
\date{}

\begin{document}
\maketitle
\tableofcontents
\newpage

%──────────────────────────────────────────────────────────────────────────────
\section{Purpose and Scope}

Audits the Beta factor: exponentially-weighted OLS slope of stock
excess returns on the ESTU cap-weighted excess benchmark over a trailing 252-day
window with a 63-day half-life (USE4 Empirical Notes, Appendix~A). \texttt{alpha}
and \texttt{r2} are retained for audit.

\smallskip
\textbf{Scope.} This audit covers every pipeline stage of
\texttt{beta\_build.ipynb} from the lab's reference run of 2026-06-11: all printed diagnostics, the
validation battery, and the output parquet schema. All numerical values are
verbatim from the executed notebook. The notebook generates no figures.

\NOTE{Reference numbers: every value in this document comes verbatim from the lab's reference run of 2026-06-11 (Sharadar data through 2026-06-10). Your rebuild will differ in row counts and tail statistics with data vintage; distribution shapes, stability coefficients, and check outcomes should match. Treat persistent sign or order-of-magnitude differences as bugs in your build until proven otherwise.}


%──────────────────────────────────────────────────────────────────────────────
\section{Output Schema}

The deliverable is \path{data/out/beta_use4.parquet}, written with
\texttt{compression="zstd"}, \newline\texttt{statistics=True}.

\begin{table}[h!]
\centering
\caption{Output schema}
\begin{tabular}{llp{7.4cm}}
\toprule
\textbf{Column} & \textbf{Dtype} & \textbf{Description} \\
\midrule
\texttt{permaticker} & \texttt{Int64} & Sharadar permanent ticker ID \\
\texttt{signal\_date} & \texttt{Date} & End-of-month rebalance date \\
\texttt{in\_estu} & \texttt{Bool} & ESTU membership \\
\texttt{mcap} & \texttt{Float64} & Market cap on signal date \\
\texttt{beta} & \texttt{Float64} & Raw descriptor: EW-OLS slope, post-trim \\
\texttt{alpha} & \texttt{Float64} & Regression intercept (audit column) \\
\texttt{beta\_score} & \texttt{Float64} & Final standardised exposure \\
\texttt{n\_obs} & \texttt{UInt32} & Valid observations in the window ($\geq 63$) \\
\texttt{r2} & \texttt{Float64} & Regression $R^2$ (audit column) \\
\bottomrule
\end{tabular}
\end{table}

%──────────────────────────────────────────────────────────────────────────────
\section{Shared Infrastructure}

All factor builds consume the shared estimation universe
\path{data/out/estu_monthly.parquet} (audited separately in
\texttt{estu\_audit.tex}); the time-series builds additionally consume the
shared daily-returns artifact (audited in \texttt{daily\_panel\_audit.tex}).
On this run the factor-level view of the shared inputs was:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item Signal calendar: \textbf{330 dates} (1999-01-29 $\to$ 2026-06-10)
  \item ESTU size: mean \textbf{2{,}494}, min \textbf{1{,}791}, max \textbf{3{,}045}
  \item Daily artifact: \textbf{38{,}997{,}999} rows, \textbf{17{,}413} tickers (1998-01-02 $\to$ 2026-06-10); loader mode \textbf{FAST} (commit-aware gate)
  \item \texttt{\_td\_to\_sig}: \textbf{7{,}153} trading days mapped; missing-benchmark dates \textbf{0} (target $<$ 30)
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Known Deviations from PDF Spec}

This section mirrors the complete NOT-IN-PDF master list from
\texttt{beta\_spec.ipynb} (\S13.) --- every judgment call the USE4 PDFs do not
specify, with the chosen solution and its revisit trigger. Rejected alternatives
are documented in the guide; the four-part validity arguments (USE4 requirement
$\to$ deviation $\to$ solution $\to$ defence) live in the pipeline-level
register, \path{docs/deviations_register.tex}.

\begin{longtable}{@{}p{0.5cm}p{3.6cm}p{6.0cm}p{3.8cm}@{}}
\caption{NOT-IN-PDF decisions (mirrors the guide master list)}\\
\toprule
\textbf{\#} & \textbf{Decision} & \textbf{Chosen solution} & \textbf{Revisit when} \\
\midrule
\endhead
1 & ESTU construction (MSCI USA IMI is proprietary) & Shared \texttt{estu\_monthly}: top 3000, ATVR screen, 3000/3500 buffer & If ESTU churn > 5\%/month \\[3pt]
2 & Risk-free rate for excess returns & Ken French daily RF, subtracted in the shared daily artifact & If validating against USE4 published betas \\[3pt]
3 & Min observations for beta & 63 ($\sim$3 months) & If coverage too low at start of sample \\[3pt]
4 & Outlier treatment & 3-sigma trim per Section 2.2 (optional) & If validation shows extreme beta values dominating cross-section \\[3pt]
5 & ESTU buffer rule for stability & Inherited from shared ESTU (3000/3500 hysteresis) & If month-over-month stability < 0.94 \\[3pt]
6 & Daily vs monthly ESTU for benchmark & Monthly (locked at signal\_date) & Probably never \\[3pt]
7 & Float-adjusted mcap \WARN{Material} & Total mcap (Sharadar doesn't have float) & If running for institutional-grade comparison to MSCI \\[3pt]
8 & Calendar start & 1999-01-01 & If we want full sample \\[3pt]
\bottomrule
\end{longtable}

%──────────────────────────────────────────────────────────────────────────────
\section{Stage-by-Stage Audit}

%──────────────────────────────────
\subsection{Stage 1 --- Shared daily-returns artifact}

\begin{description}[style=nextline]
  \item[PDF SPEC] Excess returns and the ESTU cap-weighted benchmark (built and validated once in \texttt{daily\_panel\_build.ipynb}; see \texttt{daily\_panel\_audit.tex}).
  \item[Implementation] \texttt{factor\_lib.load\_daily\_artifacts}: \texttt{daily}, ticker map, trading days, commit-aware FAST gate.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item \textbf{38{,}997{,}999} rows, \textbf{17{,}413} tickers (1998-01-02 $\to$ 2026-06-10)
      \item Working set $\approx$1.4 GB; mode \textbf{FAST}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 2 --- Load shared ESTU}

\begin{description}[style=nextline]
  \item[PDF SPEC] USE4 ESTU = MSCI USA IMI; shared \path{estu_monthly.parquet}.
  \item[Implementation] In-ESTU members + monthly calendar; mega-cap spot checks.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item 330 dates (1999-01-29 $\to$ 2026-06-10); in-ESTU rows \textbf{822{,}918}; size mean \textbf{2{,}494}
      \item Spot: JPM/MSFT \textbf{330} months; GOOGL \textbf{261}; AAPL \textbf{257}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 4 --- Per-stock Beta (EW-OLS)}

\begin{description}[style=nextline]
  \item[PDF SPEC] 252-day window, 63-day half-life exponential weights, $w_\tau = \exp(-\ln 2\,(t-\tau)/63)$.
  \item[Implementation] Vectorised window regressions per signal date; numpy reference equivalence asserted.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Numpy-reference equivalence at 2026-05-29: max diff $\mathbf{2.66\times10^{-15}}$ (target $<10^{-10}$); median $R^2$ 0.075
      \item Raw panel: \textbf{1{,}841{,}672} rows, \textbf{327} dates, \textbf{16{,}404} tickers; duplicates \textbf{0}
      \item Cross-sectional median beta \textbf{0.852}; CW ESTU beta \textbf{1.007} (target $\approx$1.0); betas in $[0,2.5]$: \textbf{83.5\%}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 5 --- Outlier trim (3$\sigma$)}

\begin{description}[style=nextline]
  \item[PDF SPEC] Methodology \S2.2 trim.
  \item[Implementation] CW mean $\pm\,3\cdot$EW std (ESTU bounds, applied to all).
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Clipped (ESTU): \textbf{10{,}024} / 813{,}270 = \textbf{1.23\%} (target 0.5--2\%)
      \item Bounds: lo mean $-0.567$; hi mean $2.580$ (max $3.863$)
      \item Post-trim at 2026-05-29 (ESTU, $n=2{,}746$): mean 1.134, std 0.836, skew 0.742, kurt 0.700
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 6 --- Standardisation}

\begin{description}[style=nextline]
  \item[PDF SPEC] CW mean 0 (ESTU), EW std 1 (ESTU), applied to all stocks.
  \item[Implementation] Standard shared machinery; ticker spot prints retained.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item CW mean: \textbf{$-$0.000000}; EW std: \textbf{1.000000}
      \item Fraction in $[\pm3]$ at 2026-05-29: \textbf{97.4\%}
      \item Spot scores: GOOGL $+0.352$, MSFT $-0.270$, AAPL $-0.294$, JPM $-0.169$, NEE $-1.133$, KO $-1.404$
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 7 --- Save}

\begin{description}[style=nextline]
  \item[PDF SPEC] Parquet deliverable with asserted schema (9 columns).
  \item[Implementation] zstd; dtype + read-back assertions.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Rows \textbf{1{,}841{,}672}; dates \textbf{327} (1999-04-30 $\to$ 2026-06-10); tickers \textbf{16{,}404}; \textbf{63.4 MB}
      \item Schema check \PASS{PASSED}; read-back \PASS{PASSED}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────────────────────────────────────────────────
\section{Validation Checks}

All checks passed on the run dated 2026-06-11. Coverage (check 3) is evaluated
on completed months only --- the final in-progress signal date is excluded
(shared convention across all factor builds).

\begin{table}[h!]
\centering
\caption{Validation results --- run date 2026-06-11}
\small
\begin{tabular}{clll}
\toprule
\textbf{Check} & \textbf{Description} & \textbf{Observed} & \textbf{Target / Status} \\
\midrule
1a & Max $|\mu_{\text{CW}}(\texttt{beta\_score})|$ & $8.39\times10^{-16}$ & $<10^{-6}$ \quad \PASS{PASS} \\
1b & Mean EW std & $1.0000$ & $1.0\pm0.02$ \quad \PASS{PASS} \\
2 & Max $|\mu_{\text{CW}}(\text{raw beta})-1|$ & $0.1629$ & $<0.20$ \quad \PASS{PASS} \\
3 & Min coverage post-2005 (completed months) & $4{,}854$ & $\geq4{,}000$ \quad \PASS{PASS} \\
4 & MoM Spearman $\rho$ mean & $0.9458$ & $>0.94$ \quad \PASS{PASS} \\
5 & Q5$-$Q1 vs ESTU mkt corr & $0.7475$ & $>0.70$ \quad \PASS{PASS} \\
6 & Disk vs memory max diff & $0.00\times10^{0}$ & $<10^{-10}$ \quad \PASS{PASS} \\
\bottomrule
\end{tabular}
\end{table}

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item \textbf{Check 2:} worst-date deviation 0.1629 at 2001-02-28 (ESTU $n=2{,}081$); mean deviation 0.0180.
  \item \textbf{Check 4:} $\rho > 0.94$ on 71.8\% of months; $> 0.90$ on 87.1\% (USE4 Table 4.2 reports 0.96--0.97).
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Spot Check at 2026-05-29}

Aggregate diagnostics at the penultimate signal date:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item \textbf{5{,}204} tickers matched in the equivalence check; alpha/r2 non-null for all
  \item Post-trim ESTU ($n=2{,}746$): mean \textbf{1.134}, std \textbf{0.836}, skew \textbf{0.742}, excess kurtosis \textbf{0.700}
\end{itemize}

\NOTE{No per-ticker factor values were printed by the build notebook
beyond the six named tickers in Stage 6; the spot check reports aggregate statistics only.}

\end{document}
