\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[table]{xcolor}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{url}
\usepackage[hidelinks,bookmarks=true,bookmarksopen=true]{hyperref}
\usepackage{siunitx}
\usepackage{parskip}

\sisetup{group-separator={,}, detect-weight=true, detect-family=true}
\renewcommand{\arraystretch}{1.15}

\definecolor{pass}{RGB}{0,140,0}
\definecolor{warn}{RGB}{200,120,0}
\definecolor{fail}{RGB}{180,0,0}
\definecolor{note}{RGB}{60,60,180}
\definecolor{accent}{HTML}{01696F}
\definecolor{ink}{HTML}{28251D}
\definecolor{muted}{HTML}{7A7974}
\definecolor{rule}{HTML}{D4D1CA}

\newcommand{\PASS}[1]{\textcolor{pass}{\textbf{#1}}}
\newcommand{\WARN}[1]{\textcolor{warn}{\textbf{#1}}}
\newcommand{\FAIL}[1]{\textcolor{fail}{\textbf{#1}}}
\newcommand{\NOTE}[1]{\textcolor{note}{\textit{#1}}}

\titleformat{\section}
  {\normalfont\large\bfseries\color{accent}}{}{0pt}{}
  [\vspace{-2pt}\textcolor{rule}{\rule{\linewidth}{0.4pt}}]

\titleformat{\subsection}
  {\normalfont\normalsize\bfseries\color{ink}}{}{0pt}{}

\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{10pt}{2pt}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyfoot[L]{\footnotesize\color{muted}use4-learning-lab --- reference run 2026-06-11}
\fancyfoot[C]{\footnotesize\color{muted}\leftmark}
\fancyfoot[R]{\footnotesize\color{muted}\thepage}

\title{\textbf{Non-Linear Beta (NLB) --- Factor Construction Audit}\\[2pt]
\large USE4-Faithful Implementation --- \texttt{nlb\_build.ipynb}\\[4pt]
\normalsize USE4 Learning Lab --- Reference Audit}
\author{Audit Date: 2026-06-11}
\date{}

\begin{document}
\maketitle
\tableofcontents
\newpage

%──────────────────────────────────────────────────────────────────────────────
\section{Purpose and Scope}

Audits the Non-Linear Beta factor: the cube of the standardised
Beta exposure, orthogonalised to Beta on a $\sqrt{\text{mcap}}$-weighted basis,
trimmed and standardised (USE4 Empirical Notes, Appendix~A). Beta inputs are read
from \path{beta_use4.parquet} so NLB is orthogonalised to the same Beta factor
that enters the risk model.

\smallskip
\textbf{Scope.} This audit covers every pipeline stage of
\texttt{nlb\_build.ipynb} from the lab's reference run of 2026-06-11: all printed diagnostics, the
validation battery, and the output parquet schema. All numerical values are
verbatim from the executed notebook. The notebook generates no figures.

\NOTE{Reference numbers: every value in this document comes verbatim from the lab's reference run of 2026-06-11 (Sharadar data through 2026-06-10). Your rebuild will differ in row counts and tail statistics with data vintage; distribution shapes, stability coefficients, and check outcomes should match. Treat persistent sign or order-of-magnitude differences as bugs in your build until proven otherwise.}


%──────────────────────────────────────────────────────────────────────────────
\section{Output Schema}

The deliverable is \path{data/out/nlb_use4.parquet}, written with
\texttt{compression="zstd"}, \newline\texttt{statistics=True}.

\begin{table}[h!]
\centering
\caption{Output schema}
\begin{tabular}{llp{7.4cm}}
\toprule
\textbf{Column} & \textbf{Dtype} & \textbf{Description} \\
\midrule
\texttt{permaticker} & \texttt{Int64} & Sharadar permanent ticker ID \\
\texttt{signal\_date} & \texttt{Date} & End-of-month rebalance date \\
\texttt{in\_estu} & \texttt{Bool} & ESTU membership \\
\texttt{beta} & \texttt{Float64} & Raw beta from the Beta deliverable \\
\texttt{beta\_z} & \texttt{Float64} & Standardised Beta (\texttt{beta\_score} upstream) \\
\texttt{beta\_z\_cubed} & \texttt{Float64} & $\texttt{beta\_z}^{3}$ \\
\texttt{nlb\_raw} & \texttt{Float64} & Orthogonalisation residual (trimmed in Stage 5) \\
\texttt{nlb\_score} & \texttt{Float64} & Final standardised exposure \\
\texttt{n\_obs} & \texttt{UInt32} & Inherited from the Beta deliverable \\
\texttt{mcap} & \texttt{Float64} & Market cap on signal date \\
\bottomrule
\end{tabular}
\end{table}

%──────────────────────────────────────────────────────────────────────────────
\section{Shared Infrastructure}

All factor builds consume the shared estimation universe
\path{data/out/estu_monthly.parquet} (audited separately in
\texttt{estu\_audit.tex}); the time-series builds additionally consume the
shared daily-returns artifact (audited in \texttt{daily\_panel\_audit.tex}).
On this run the factor-level view of the shared inputs was:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item Signal calendar: \textbf{330 dates} (1999-01-29 $\to$ 2026-06-10)
  \item ESTU size: mean \textbf{2{,}494}, min \textbf{1{,}791}, max \textbf{3{,}045}
  \item Daily artifact: \textbf{38{,}997{,}999} rows, \textbf{17{,}413} tickers (1998-01-02 $\to$ 2026-06-10); loader mode \textbf{FAST} (commit-aware gate)
  \item \texttt{\_td\_to\_sig}: \textbf{7{,}153} trading days mapped; missing-benchmark dates \textbf{0} (target $<$ 30)
  \item Beta artifact: \textbf{1{,}841{,}672} rows, \textbf{327} dates, \textbf{16{,}404} tickers
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Known Deviations from PDF Spec}

This section mirrors the complete NOT-IN-PDF master list from
\texttt{nlb\_spec.ipynb} (\S16.) --- every judgment call the USE4 PDFs do not
specify, with the chosen solution and its revisit trigger. Rejected alternatives
are documented in the guide; the four-part validity arguments (USE4 requirement
$\to$ deviation $\to$ solution $\to$ defence) live in the pipeline-level
register, \path{docs/deviations_register.tex}.

\begin{longtable}{@{}p{0.5cm}p{3.6cm}p{6.0cm}p{3.8cm}@{}}
\caption{NOT-IN-PDF decisions (mirrors the guide master list)}\\
\toprule
\textbf{\#} & \textbf{Decision} & \textbf{Chosen solution} & \textbf{Revisit when} \\
\midrule
\endhead
1 & ESTU construction (MSCI USA IMI is proprietary) & Shared \texttt{estu\_monthly}: top 3000, ATVR screen, 3000/3500 buffer & If ESTU churn > 5\%/month \\[3pt]
2 & Risk-free rate for excess returns & Ken French daily RF, subtracted in the shared daily artifact & If validating against USE4 published betas \\[3pt]
3 & Min observations for beta & 63 ($\sim$3 months) & If coverage too low at start of sample \\[3pt]
4 & Outlier treatment & 3$\sigma$ trim (family standard); ESTU clip 2.42\% documented & If ESTU clip rate exceeds $\sim$3\% \\[3pt]
5 & Pairwise vs joint orthogonalization & Pairwise (Beta only) per Appendix A literal reading & When building the multi-factor risk model --- orthogonalize jointly in the full CSR \\[3pt]
6 & ESTU buffer rule for stability & Inherited from shared ESTU (3000/3500 hysteresis) & If month-over-month exposure stability < 0.85 \\[3pt]
7 & Daily vs monthly ESTU membership for benchmark & Monthly (locked at signal\_date) & Probably never --- monthly matches index practice \\[3pt]
8 & Treatment of stocks delisted mid-window & Include observations up to delisting & Edge case; coverage check will flag if material \\[3pt]
9 & Float-adjusted mcap \WARN{Material} & Total mcap (Sharadar doesn't have float) & If running for institutional-grade comparison to MSCI \\[3pt]
10 & Calendar start & 1999-01-01 & If we want full sample \\[3pt]
\bottomrule
\end{longtable}

%──────────────────────────────────────────────────────────────────────────────
\section{Stage-by-Stage Audit}

%──────────────────────────────────
\subsection{Stage 1 --- Shared daily-returns artifact}

\begin{description}[style=nextline]
  \item[PDF SPEC] Excess returns and the ESTU cap-weighted benchmark (built and validated once in \texttt{daily\_panel\_build.ipynb}; see \texttt{daily\_panel\_audit.tex}).
  \item[Implementation] \texttt{factor\_lib.load\_daily\_artifacts}: \texttt{daily}, ticker map, trading days, commit-aware FAST gate.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item \textbf{38{,}997{,}999} rows, \textbf{17{,}413} tickers (1998-01-02 $\to$ 2026-06-10)
      \item Working set $\approx$1.4 GB; mode \textbf{FAST}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 2 --- Load shared ESTU}

\begin{description}[style=nextline]
  \item[PDF SPEC] USE4 ESTU = MSCI USA IMI; shared \path{estu_monthly.parquet}.
  \item[Implementation] In-ESTU members + monthly calendar; mega-cap spot checks.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item 330 dates (1999-01-29 $\to$ 2026-06-10); in-ESTU rows \textbf{822{,}918}; size mean \textbf{2{,}494}
      \item Spot: JPM/MSFT \textbf{330} months; GOOGL \textbf{261}; AAPL \textbf{257}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 4--7 --- Cube + orthogonalise}

\begin{description}[style=nextline]
  \item[PDF SPEC] Cube the standardised Beta; orthogonalise to Beta on a regression-weighted basis.
  \item[Implementation] \texttt{beta\_z} $=$ upstream \texttt{beta\_score}; cube; per-date $\sqrt{\text{mcap}}$ WLS of the cube on beta, ESTU fit, residual applied to all.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Beta artifact loaded: \textbf{1{,}841{,}672} rows, 327 dates, 16{,}404 tickers
      \item Pre-trim orthogonality: $\mathbf{3.40\times10^{-15}}$ (target $<10^{-10}$)
      \item Panel: \textbf{1{,}841{,}672} rows; duplicates \textbf{0}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 8 --- Outlier trim (3$\sigma$)}

\begin{description}[style=nextline]
  \item[PDF SPEC] ``the factor is winsorized'' (method unspecified).
  \item[Implementation] CW mean $\pm\,3\cdot$EW std (ESTU bounds, applied to all).
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Clipped (ESTU): \textbf{19{,}719} / 813{,}270 = \textbf{2.42\%} (target 0.5--2\%)
      \item Bounds: lo mean $-9.411$; hi mean $9.309$ (max $13.394$)
      \item Post-trim at 2026-05-29 (ESTU, $n=2{,}746$): mean $-0.128$, std 2.215, skew 1.203, kurt 3.931
    \end{itemize}
  \item[Deviation] \WARN{ESTU clip rate 2.42\% is 0.42pp above the 0.5--2\% band --- cubic distributions have structurally heavy tails. Accepted; the trim still reduces post-trim skew/kurtosis dramatically vs the raw cube.}
  \item[Status] \WARN{PARTIAL MATCH} (clip rate above band, documented)
\end{description}

%──────────────────────────────────
\subsection{Stage 9 --- Standardisation}

\begin{description}[style=nextline]
  \item[PDF SPEC] CW mean 0 (ESTU), EW std 1 (ESTU), applied to all stocks.
  \item[Implementation] Standard shared machinery.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item CW mean: \textbf{0.000000}; EW std: \textbf{1.000000}
      \item Fraction in $[\pm3]$: all \textbf{93.8\%} / ESTU \textbf{97.0\%}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 10 --- Save}

\begin{description}[style=nextline]
  \item[PDF SPEC] Parquet deliverable with asserted schema (10 columns).
  \item[Implementation] zstd; dtype + read-back assertions.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Rows \textbf{1{,}841{,}672}; dates \textbf{327} (1999-04-30 $\to$ 2026-06-10); tickers \textbf{16{,}404}; \textbf{76.1 MB}
      \item Schema check \PASS{PASSED}; read-back \PASS{PASSED}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────────────────────────────────────────────────
\section{Validation Checks}

All checks passed on the run dated 2026-06-11. Coverage (check 3) is evaluated
on completed months only --- the final in-progress signal date is excluded
(shared convention across all factor builds).

\begin{table}[h!]
\centering
\caption{Validation results --- run date 2026-06-11}
\small
\begin{tabular}{clll}
\toprule
\textbf{Check} & \textbf{Description} & \textbf{Observed} & \textbf{Target / Status} \\
\midrule
1a & Max $|\mu_{\text{CW}}(\texttt{nlb\_score})|$ & $6.97\times10^{-17}$ & $<10^{-6}$ \quad \PASS{PASS} \\
1b & Mean EW std & $1.0000$ & $1.0\pm0.02$ \quad \PASS{PASS} \\
2 & $\sqrt{\text{mcap}}$-wtd orthogonality (pre-trim) & $3.40\times10^{-15}$ & $<10^{-10}$ \quad \PASS{PASS} \\
3 & Min coverage post-2005 (completed months) & $4{,}854$ & $\geq4{,}000$ \quad \PASS{PASS} \\
4 & MoM Spearman $\rho$ mean & $0.8723$ & $>0.85$ \quad \PASS{PASS} \\
5 & Q5$-$Q1 vs ESTU mkt corr & $0.6597$ & $>0.30$ \quad \PASS{PASS} \\
6 & Disk vs memory max diff & $0.00\times10^{0}$ & $<10^{-10}$ \quad \PASS{PASS} \\
\bottomrule
\end{tabular}
\end{table}

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item \textbf{Check 2} is evaluated pre-trim: the trim deliberately moves extreme residuals, so post-trim orthogonality is neither expected nor checked.
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Spot Check at 2026-05-29}

Aggregate diagnostics at the penultimate signal date:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item Post-trim ESTU ($n=2{,}746$): mean \textbf{$-$0.128}, std \textbf{2.215}, skew \textbf{1.203}, excess kurtosis \textbf{3.931}
\end{itemize}

\NOTE{No per-ticker factor values were printed by the build notebook
; the spot check reports aggregate statistics only.}

\end{document}
