\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[table]{xcolor}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{url}
\usepackage[hidelinks,bookmarks=true,bookmarksopen=true]{hyperref}
\usepackage{siunitx}
\usepackage{parskip}

\sisetup{group-separator={,}, detect-weight=true, detect-family=true}
\renewcommand{\arraystretch}{1.15}

\definecolor{pass}{RGB}{0,140,0}
\definecolor{warn}{RGB}{200,120,0}
\definecolor{fail}{RGB}{180,0,0}
\definecolor{note}{RGB}{60,60,180}
\definecolor{accent}{HTML}{01696F}
\definecolor{ink}{HTML}{28251D}
\definecolor{muted}{HTML}{7A7974}
\definecolor{rule}{HTML}{D4D1CA}

\newcommand{\PASS}[1]{\textcolor{pass}{\textbf{#1}}}
\newcommand{\WARN}[1]{\textcolor{warn}{\textbf{#1}}}
\newcommand{\FAIL}[1]{\textcolor{fail}{\textbf{#1}}}
\newcommand{\NOTE}[1]{\textcolor{note}{\textit{#1}}}

\titleformat{\section}
  {\normalfont\large\bfseries\color{accent}}{}{0pt}{}
  [\vspace{-2pt}\textcolor{rule}{\rule{\linewidth}{0.4pt}}]

\titleformat{\subsection}
  {\normalfont\normalsize\bfseries\color{ink}}{}{0pt}{}

\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{10pt}{2pt}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyfoot[L]{\footnotesize\color{muted}use4-learning-lab --- reference run 2026-06-11}
\fancyfoot[C]{\footnotesize\color{muted}\leftmark}
\fancyfoot[R]{\footnotesize\color{muted}\thepage}

\title{\textbf{Non-Linear Size (NLS) --- Factor Construction Audit}\\[2pt]
\large USE4-Faithful Implementation --- \texttt{nls\_build.ipynb}\\[4pt]
\normalsize USE4 Learning Lab --- Reference Audit}
\author{Audit Date: 2026-06-11}
\date{}

\begin{document}
\maketitle
\tableofcontents
\newpage

%──────────────────────────────────────────────────────────────────────────────
\section{Purpose and Scope}

Audits the Non-Linear Size factor: the cube of the standardised
Size exposure, orthogonalised to Size on a $\sqrt{\text{mcap}}$-weighted basis,
winsorised (1\%/99\%) and standardised (USE4 Empirical Notes, Appendix~A, p.~55).
Size inputs are read from \path{size_use4.parquet} (\texttt{log\_mcap} $:=$
post-trim \texttt{lncap}, \texttt{size\_z} $:=$ \texttt{size\_score}), so NLS is
orthogonalised to the same Size factor that enters the risk model.

\smallskip
\textbf{Scope.} This audit covers every pipeline stage of
\texttt{nls\_build.ipynb} from the lab's reference run of 2026-06-11: all printed diagnostics, the
validation battery, and the output parquet schema. All numerical values are
verbatim from the executed notebook. The notebook generates no figures.

\NOTE{Reference numbers: every value in this document comes verbatim from the lab's reference run of 2026-06-11 (Sharadar data through 2026-06-10). Your rebuild will differ in row counts and tail statistics with data vintage; distribution shapes, stability coefficients, and check outcomes should match. Treat persistent sign or order-of-magnitude differences as bugs in your build until proven otherwise.}


%──────────────────────────────────────────────────────────────────────────────
\section{Output Schema}

The deliverable is \path{data/out/nls_use4.parquet}, written with
\texttt{compression="zstd"}, \newline\texttt{statistics=True}.

\begin{table}[h!]
\centering
\caption{Output schema}
\begin{tabular}{llp{7.4cm}}
\toprule
\textbf{Column} & \textbf{Dtype} & \textbf{Description} \\
\midrule
\texttt{permaticker} & \texttt{Int64} & Sharadar permanent ticker ID \\
\texttt{signal\_date} & \texttt{Date} & End-of-month rebalance date \\
\texttt{in\_estu} & \texttt{Bool} & ESTU membership \\
\texttt{mcap} & \texttt{Float64} & Total market cap on signal date (dollars) \\
\texttt{log\_mcap} & \texttt{Float64} & \texttt{lncap} from the SIZE deliverable (post-trim) \\
\texttt{size\_z} & \texttt{Float64} & \texttt{size\_score} from the SIZE deliverable \\
\texttt{size\_z\_cubed} & \texttt{Float64} & $\texttt{size\_z}^{3}$ \\
\texttt{nls\_raw} & \texttt{Float64} & Orthogonalisation residual (winsorised in Stage 5) \\
\texttt{nls\_score} & \texttt{Float64} & Final standardised exposure \\
\bottomrule
\end{tabular}
\end{table}

%──────────────────────────────────────────────────────────────────────────────
\section{Shared Infrastructure}

All factor builds consume the shared estimation universe
\path{data/out/estu_monthly.parquet} (audited separately in
\texttt{estu\_audit.tex}); the time-series builds additionally consume the
shared daily-returns artifact (audited in \texttt{daily\_panel\_audit.tex}).
On this run the factor-level view of the shared inputs was:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item Signal calendar: \textbf{330 dates} (1999-01-29 $\to$ 2026-06-10)
  \item ESTU size: mean \textbf{2{,}494}, min \textbf{1{,}791}, max \textbf{3{,}045}
  \item Daily artifact: \textbf{38{,}997{,}999} rows, \textbf{17{,}413} tickers (1998-01-02 $\to$ 2026-06-10); loader mode \textbf{FAST} (commit-aware gate)
  \item \texttt{\_td\_to\_sig}: \textbf{7{,}153} trading days mapped; missing-benchmark dates \textbf{0} (target $<$ 30)
  \item SIZE artifact: \textbf{1{,}773{,}820} rows, \textbf{330} dates, \textbf{16{,}829} tickers
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Known Deviations from PDF Spec}

This section mirrors the complete NOT-IN-PDF master list from
\texttt{nls\_spec.ipynb} (\S15.) --- every judgment call the USE4 PDFs do not
specify, with the chosen solution and its revisit trigger. Rejected alternatives
are documented in the guide; the four-part validity arguments (USE4 requirement
$\to$ deviation $\to$ solution $\to$ defence) live in the pipeline-level
register, \path{docs/deviations_register.tex}.

\begin{longtable}{@{}p{0.5cm}p{3.6cm}p{6.0cm}p{3.8cm}@{}}
\caption{NOT-IN-PDF decisions (mirrors the guide master list)}\\
\toprule
\textbf{\#} & \textbf{Decision} & \textbf{Chosen solution} & \textbf{Revisit when} \\
\midrule
\endhead
1 & ESTU construction (MSCI USA IMI is proprietary) & Shared \texttt{estu\_monthly}: top 3000, ATVR screen, 3000/3500 buffer & If ESTU churn > 5\%/month \\[3pt]
2 & Winsorization percentiles & 1\% / 99\% & If validation shows fat tails after winsorization \\[3pt]
3 & Pairwise vs joint orthogonalization & Pairwise (Size only) per Appendix A literal reading & When building the multi-factor risk model --- orthogonalize jointly in the full CSR \\[3pt]
4 & ESTU buffer rule for stability & None (re-rank monthly) & If month-over-month exposure stability < 0.95 \\[3pt]
5 & Float-adjusted mcap \WARN{Material} & Total mcap (Sharadar doesn't have float) & If running for institutional-grade comparison to MSCI \\[3pt]
6 & Sharadar \texttt{marketcap} unit (dollars vs millions) & Dollars (ln values $\approx$ 18--23); documented in code and audits & Document the unit in the code; downstream consumers may care \\[3pt]
7 & Calendar start & 1999-01-01 (match NLB) & Probably never --- match NLB for consistency \\[3pt]
8 & Regress cube on \texttt{log\_mcap} or on \texttt{size\_z} & Either --- residuals are identical & Pure stylistic choice \\[3pt]
\bottomrule
\end{longtable}

%──────────────────────────────────────────────────────────────────────────────
\section{Stage-by-Stage Audit}

%──────────────────────────────────
\subsection{Stage 1 --- Shared daily-returns artifact}

\begin{description}[style=nextline]
  \item[PDF SPEC] Excess returns and the ESTU cap-weighted benchmark (built and validated once in \texttt{daily\_panel\_build.ipynb}; see \texttt{daily\_panel\_audit.tex}).
  \item[Implementation] \texttt{factor\_lib.load\_daily\_artifacts}: \texttt{daily}, ticker map, trading days, commit-aware FAST gate.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item \textbf{38{,}997{,}999} rows, \textbf{17{,}413} tickers (1998-01-02 $\to$ 2026-06-10)
      \item Working set $\approx$1.4 GB; mode \textbf{FAST}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 2 --- Load shared ESTU}

\begin{description}[style=nextline]
  \item[PDF SPEC] USE4 ESTU = MSCI USA IMI; shared \path{estu_monthly.parquet}.
  \item[Implementation] In-ESTU members + monthly calendar; mega-cap spot checks.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item 330 dates (1999-01-29 $\to$ 2026-06-10); in-ESTU rows \textbf{822{,}918}; size mean \textbf{2{,}494}
      \item Spot: JPM/MSFT \textbf{330} months; GOOGL \textbf{261}; AAPL \textbf{257}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 3--6 --- Size input, cube, orthogonalise}

\begin{description}[style=nextline]
  \item[PDF SPEC] ``First, the standardized Size exposure is cubed. The resulting factor is then orthogonalized to the Size factor on a regression-weighted basis.''
  \item[Implementation] \texttt{log\_mcap}/\texttt{size\_z} from \path{size_use4.parquet}; cube; per-date $\sqrt{\text{mcap}}$ WLS, ESTU fit, residual applied to all.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item SIZE artifact: \textbf{1{,}773{,}820} rows; coverage after mcap guard: \textbf{1{,}730{,}893} rows, \textbf{16{,}475} tickers
      \item log\_mcap median \textbf{19.868} (dollars; $\ln(4\times10^{8})\approx19.9$)
      \item Pre-winsor orthogonality: $\mathbf{8.00\times10^{-13}}$ (target $<10^{-10}$); duplicates \textbf{0}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 7 --- Winsorisation (1\%/99\%)}

\begin{description}[style=nextline]
  \item[PDF SPEC] ``the factor is winsorized''.
  \item[Implementation] ESTU 1\%/99\% quantiles per date, clip applied to all stocks ($\sigma$-based trim produces near-zero clipping on the skewed cubic).
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Clipped (ESTU): \textbf{16{,}421} / 822{,}918 = \textbf{2.00\%} (by construction)
      \item Clipped (all universe): \textbf{782{,}660} --- intentional: ESTU bounds pin non-ESTU micro-cap cubes
      \item Bounds: lo mean $-20.334$; hi mean $5.365$
      \item Post-winsor at 2026-05-29 (ESTU, $n=2{,}761$): mean $-4.288$, std 11.525, skew $-0.983$, kurt $-0.239$
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 8 --- Standardisation}

\begin{description}[style=nextline]
  \item[PDF SPEC] Winsorise \emph{then} standardise; CW mean 0, EW std 1 (ESTU), applied to all.
  \item[Implementation] Standard shared machinery.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item CW mean: \textbf{0.000000}; EW std: \textbf{1.000000}
      \item Fraction in $[\pm3]$: all \textbf{100.0\%} / ESTU \textbf{100.0\%} (signature of winsorise-then-standardise)
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────
\subsection{Stage 9 --- Save}

\begin{description}[style=nextline]
  \item[PDF SPEC] Parquet deliverable with asserted schema (9 columns, no n\_obs).
  \item[Implementation] zstd; dtype + read-back assertions.
  \item[Observed]\phantom{x}
    \begin{itemize}[nosep, topsep=-\parskip]
      \item Rows \textbf{1{,}730{,}893}; dates \textbf{330}; tickers \textbf{16{,}475}; \textbf{52.2 MB}
      \item Schema check \PASS{PASSED}; read-back \PASS{PASSED}
    \end{itemize}
  \item[Status] \PASS{MATCHES SPEC}
\end{description}

%──────────────────────────────────────────────────────────────────────────────
\section{Validation Checks}

All checks passed on the run dated 2026-06-11. Coverage (check 3) is evaluated
on completed months only --- the final in-progress signal date is excluded
(shared convention across all factor builds).

\begin{table}[h!]
\centering
\caption{Validation results --- run date 2026-06-11}
\small
\begin{tabular}{clll}
\toprule
\textbf{Check} & \textbf{Description} & \textbf{Observed} & \textbf{Target / Status} \\
\midrule
1a & Max $|\mu_{\text{CW}}(\texttt{nls\_score})|$ & $7.00\times10^{-17}$ & $<10^{-6}$ \quad \PASS{PASS} \\
1b & Mean EW std & $1.0000$ & $1.0\pm0.02$ \quad \PASS{PASS} \\
2 & $\sqrt{\text{mcap}}$-wtd orthogonality (pre-winsor) & $8.00\times10^{-13}$ & $<10^{-10}$ \quad \PASS{PASS} \\
3 & Min coverage post-2005 (completed months) & $4{,}578$ & $\geq4{,}000$ \quad \PASS{PASS} \\
4 & MoM Spearman $\rho$ mean (ESTU) & $0.9882$ & $>0.95$ \quad \PASS{PASS} \\
5 & Q5$-$Q1 vs ESTU mkt $|\text{corr}|$ & $-0.1331$ & $|\text{corr}|<0.20$ \quad \PASS{PASS} \\
6 & Disk vs memory max diff & $0.00\times10^{0}$ & $<10^{-10}$ \quad \PASS{PASS} \\
\bottomrule
\end{tabular}
\end{table}

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item \textbf{Check 2} is pre-winsorisation only --- the winsorise step deliberately breaks the weighted inner product (clipped stocks have extreme log\_mcap).
  \item \textbf{Check 4:} 0.9882 sits inside the 0.98--0.99 range USE4 Table 4.2 reports for NLS.
\end{itemize}

%──────────────────────────────────────────────────────────────────────────────
\section{Spot Check at 2026-05-29}

Aggregate diagnostics at the penultimate signal date:

\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
  \item Post-winsor ESTU ($n=2{,}761$): mean \textbf{$-$4.288}, std \textbf{11.525}, skew \textbf{$-$0.983}, excess kurtosis \textbf{$-$0.239}
\end{itemize}

\NOTE{No per-ticker factor values were printed by the build notebook
; the spot check reports aggregate statistics only.}

\end{document}
